A Hitchhiker's Guide to Reproducible Research" # replace with something serious

### 2021-04-06 manuscript.Rmd
fix typo

---
title: "A Hitchhiker's Guide to Reproducible Research" # replace with something serious
output: bookdown::pdf_document2
repro:
  packages:
    - usethis
    - aaronpeikert/repro@7bfaf98
    - here
    - rstudio/webshot2@f62e743
csl: apa7.csl
numbered_headings: true
toc_depth: 5
bibliography: references.bib
---

@@

<!-- define reproducibility -->
<!-- is a direct quote from my master thesis -->
<!-- I wouldn't want to cite it. @brandmaier, would you please rewrite? -->
A reproducible scientific product allows other researchers to obtain the same results from the same dataset in a way that enables substantive criticism and therefore facilitates replication.
<!-- lure the reader in -->
The research community has long accepted reproducibility as part of "good scientific practice", but has not has not benefited much from this self-imposed requirement.
<!-- expand: who is calling it a requirement, maybe metascientific arguments-->
Reproducing an empirical study is often tedious and frustrating work for the researcher.
However, if reproduction would require mere minutes instead of hours, it could greatly facilitate collaboration within and across research projects.
<!-- Guiding principle: spending machine compute time instead of human research time -->
To archive this boost, we must place the burden of reproducing something upon computers instead of human researchers.
A scientific document should therefore be understood by researchers but reproduced by computers.

This tutorial walks through the creation of such a document that can be reproduced automatically.
Though reproducibility is but a small part of the research process, it can serve as the mortar of open science building blocks (e.g. preregistration, open data, postpublication review).
Therefore, we show the whole lifecycle of an open research project, where some parts are essential to archive reproducibility.
All code chunks are meant for the reader to try out for themselves.
Still, the reader can safely skip sections marked as optional, when unfit for their research question or methods.
However, any empirical data analysis will suffer from the following threats to reproducibility:
<!-- Which problems are solved by which tool (like in Andreas talk)-->
<!--stolen from: https://brandmaier.github.io/reproducible-data-analysis-materials/KULeuvenQuantPsy2020.html#11 -->

@@

We deem each concept necessary to archive longterm reproducibility.
However, which tools you use to implement these concepts is a matter of taste and project requirements.
This tutorial follows the recommendations of @Peikert2019 and utilize Git for version control, RMarkdown for dynamic document creation, Make for dependency tracking, and Docker for software management.
Their interplay is shown in Figure \@ref(fig:schematic).
At several points, we suggest other viable alternatives.

```{r schematic, eval = TRUE, echo = FALSE, fig.cap="Schematic illustration of the interplay of the four components (in dashed boxes) central to the reproducible workflow: version control (Git), dependency tracking (Make), software management (Docker), and dynamic document generation (R Markdown). Git tracks changes to the project over time. Make manages dependencies among the files. Docker provides a container in which the final report is built using dynamic document generation in R Markdown. Reproduced from @Peikert2019.  "}
# file gets downloaded in Makefile
knitr::include_graphics("images/nutshell.svg", auto_pdf = TRUE)
```

@@

<!-- researcher should begin early to enjoy most benefits -->
Reproducibility is invaluable for effective collaboration---including with your "past self".
While you might be tempted to consider reproducibility only when you about to finish a project, upfront investments pay of through your increased productivity.
Therefore, we suggest starting early, ideally from the moment you decide to pursue a project to incorporate tools to help you archive reproducibility gradually.
The first step is to organize all files in one folder.
This way, you never lose anything, and you can easily share the files.
<!-- Hands-on: Introducing RProjects-->
RStudio's RProjects facilitate this concept of a project folder.
To create one, click on:

@@

## Markdown

@@

```{r, results='asis', echo=FALSE}
# in case we want to reuse this example later
markdown_example <- "
<!--this is a Markdown file -->
# Header

Normal text.
Important **word**.

@@

We will later see how this simplicity enables us to create many different document formats (i.e. Word, PDF, HTML) and incorporate code (i.e. R, Python, Julia).

## Git

Another advantage of Markdown is that you can easily version-control it with Git.
You may be familiar with Microsoft Words' "Track Changes"-Feature and version control is similar in its basic idea but more potent in its features.
Instead of tracking only one file, you can keep track of the whole project directory, and instead of losing all changes when you agree to them, the entire history is preserved transparently.
For example, this paper has been going through `r length(git2r::commits())` `r link("https://github.com/aaronpeikert/repro-tutorial/commits/master", "iterations")`.
To activate Git in a given project, you can call:
<!-- Hands-on: Introduce Git -->

@@

And you could commit it in Git:
<!--I am leaning towards doing everything in R, not sure though-->
<!--For Git we maybe should use RStudio's Git pane?-->

@@

And GitHub can show you what has changed, like in Figure \@ref(fig:idea-change).
GitHub is in its essence a webinterface to Git and you may learn more about how to use it under the section on [GitHub].

```{r idea-change-screenshot, include=FALSE}
# unfortunately css selectors won't work for pdf so we'll use png instead
idea_change <- here::here("images", "idea-change.png")
webshot2::webshot("https://github.com/aaronpeikert/repro-tutorial/commit/5b3f4641ff542158184f85d458880c35e8f09f3c?diff=split",
                  file = idea_change,
                  selector = "div#diff-2f8beaa39e5c98706d8abb3361a8383776f8df8a195a21c0e6eeb25e5aa48f45", # only select file change
                  zoom = 2)# higher resolution
```

@@

<!--I am troubled by how we should recommend to learn git-->
Git can be challenging to learn, because we are used to learn alone and Git is best learned with others.
We recommend to find a partner to attend an online lecture or workshop to learn the ins and outs of Git.

@@

There are probably few things as frustrating as realizing that the data that you gathered cannot answer the research question you had in mind.
A step to prevent unpleasant surprises is to simulate data.
Simulated data lets you verify that the data you expect to gather and the analysis that you plan fit together.

<!--Hands-on: build simple functions tailored for your analysis-->

## Preregestration (optional)

@@

### R Markdown

@@

One of the greatest benefits of open research is that other people can freely reuse materials.
To use your materials, they first have to find them and learn how to use them.
Interested users find out about your project via the readme.
The readme is the first thing they see when they visit your project on [GitHub].
It should explain to them what your project is about and how to reproduce your project.
You could for example include the abstract of your (planned) paper here.

To add a readme to a repository run:

@@

This command will create a file with the name `README.Rmd`.
<!--knit/rmarkdown will be explained in section above-->
When you knit this file, it in turn creates a `README.md`.
That means you can include R code or plots etc. within the readme.
The `README.md` file will look pleasant in the browser, just like in Figure \@ref(fig:readme).

```{r readme-screenshot, include=FALSE}
# unfortunately css selectors won't work for pdf so we'll use png instead
readme <- here::here("images", "readme.png")
webshot2::webshot("https://github.com/aaronpeikert/repro-tutorial/",
                  file = readme,
                  vwidth = 1920, # height & width are not resolution
                  vheight = 1080, # but what website thinks the browser size is
                  selector = "div#readme",
                  zoom = 2) # higher resolution
```

```{r readme, echo=FALSE, fig.cap="A screenshot of how a README.md is rendered/presented on GitHub.  ", out.width='100%'}
knitr::include_graphics(readme)
```

Even when other researchers are excited to use your code or materials, they are usually legally inhibited from to doing so.
That is because in many countries (i.e. the USA, European Union) the creator of something has exclusive right to use their creation.
A license allows other people to use what you have created.
<!---exact copy from our previous manuscript, do we need to change? -->
In our experience, the [Creative Commons - Attribution license (CC-BY)](https://creativecommons.org/licenses/by/4.0/) is often appropriate for sharing texts, R Markdown files, generated figures, and other media, whereas scripts and any other computer code are often best shared under the [MIT license](https://opensource.org/licenses/MIT) (or similar permissive licenses).
Both licenses assure maximal freedom for future users while requiring the attribution of the original authors in derivative work.
These licenses are also in line with the recommendations by the Reproducible Research Standard [@stoddenEnhancingReproducibilityComputational2016; @stoddenEnablingReproducibleResearch2009].
A great resource to choose a license is [choosealicense.com](https://choosealicense.com), however, no resource, including our recommendation, replaces legal advice.
<!-- plagiary end -->
It is important to get written consent from collaborators when you change/add the license.
However, if their first contribution occurs after you have added a license, they give their consent by contributing.
To add a CC-BY license to your project simply run (there are similar commands for other licenses):

<!--Hands on: Add a license-->
```r
usethis::use_ccby_license()
```

We strongly suggest that you strive for respectful interactions with anyone who might want to contribute.
In fact, we hope it comes naturally to you.
To communicate that contributors can expect a welcoming and respectful environment, you can add the `r link( "https://www.contributor-covenant.org", "Contributor Covenant")` to your repository.
Contributing to an open project is daunting and a public pledge to value any contributions (even when you do not incorporate them) can help others to take the first step.
The Contributor Covenant is a code of conduct used by many open source project.

To add the Contributor Covenant to your project use:

@@

However, there is some controversy around how to handle violations of a code of conduct.
For that or any other reason, you may want to adapt the Contributor Covenant, to fit the needs of you project (which you are allowed to to because it is published with a CC-BY license).

If you have considered what you want to tell the public about the project with a readme, have chosen a license and thought about how to make collaboration welcoming, you may publish your Git repository.
See the section on [GitHub] on how to publish your project.

@@

Often, reproducibility is reduced to the merits it has for the reader.
However, we argue that reproducibility is vital for collaboration, not only in the sense that the consumer of a scientific product build upon it, but also that the creators work together.
The time that you invest into ensuring reproducibility, is directed at ensuring effective collaboration.
Therefore, reproducibility is at its core about collaboration.

@@

On your local machine, you use Git to track changes.
Online services like GitHub allow you to share these changes with others.
GitHub is not the only solution; there is also [GitLab](gitlab.com) or [BitBucket](https://bitbucket.org), which are as feature-rich as GitHub.
GitLab has the advantage that your institution can host it themselves; therefore, you are not relying on commercial service providers.
We focus on GitHub because it is the most popular service.
To upload your changes to GitHub, you need to create an account under https://github.com/join.
As a researcher, you are eligible for the `r link("https://help.github.com/en/articles/applying-for-an-educator-or-researcher-discount", "Researcher/Educator discount")`.
The following command creates a new GitHub repository under your GitHub account, beware this is public:
<!--Hands-on: Introduce Github-->

@@

To create a private repository (only you and people you invite have access), use:

```r
usethis::use_github(private = TRUE)
```

@@

# Analyzing Data

## Internal Dependencies

@@

## External Dependencies

@@

### renv (alternative)

### Singularity (optional)
<!-- Hands on: add data via automate() -->
<!-- Hands on: automatically download data via manually editing the Makefile -->

## Data Integrity (optional)

@@

## Anonymising Data (optional)

@@

# Peer Review (optional)

# Post Publication Review (optional)

# Archiving

@@

# Checklist

>>>>

---
title: "A Hitchhiker's Guide to Reproducible Research" # replace with something serious
output: bookdown::pdf_document2
repro:
  packages:
    - usethis
    - repro
    - here
  scripts:
    - R/lavaan_helper.R
    - R/simulation.R
    - R/link.R
csl: apa7.csl
bibliography: ["references.bib", "packages.bib"]
---

@@

<!-- define reproducibility -->
<!-- is a direct quote from my master thesis -->
<!-- I wouldn't want to cite it. @brandmaier, would you please rewrite? -->
A reproducible scientific product allows other researchers to obtain the same results from the same dataset in a way that enables substantive criticism and therefore facilitates replication.
<!-- lure the reader in -->
This notion of independent verifiability has long been part of "good scientific practice", but has not reached its full potential to benefit the scientific community.
<!-- expand: who is calling it a requirement, maybe metascientific arguments-->
Reproducing an empirical study is often tedious and frustrating work for the researcher.
However, if reproduction would require mere minutes instead of hours, it could greatly facilitate collaboration within and across research projects.
<!-- Guiding principle: spending machine compute time instead of human research time -->
To archive this boost, we must place the burden of reproducing something upon computers instead of human researchers.
A scientific document should therefore be understood by researchers but reproduced by computers.

This tutorial walks through the creation of such a document that can be reproduced automatically.
Though reproducibility is but a small part of the research process, it can serve as the mortar of open science building blocks (e.g. preregistration, open data, postpublication review).
Therefore, we show the whole lifecycle of an open research project, where some parts are essential to archive reproducibility.
Still, the reader can safely skip sections marked as optional, when unfit for their research question or methods.
However, any empirical data analysis will suffer from the following threats to reproducibility:
<!-- Which problems are solved by which tool (like in Andreas talk)-->
<!--stolen from: https://brandmaier.github.io/reproducible-data-analysis-materials/KULeuvenQuantPsy2020.html#11 -->

@@

We deem each concept necessary to archive longterm reproducibility.
However, which tools you use to implement these concepts is a matter of taste and project requirements.
This tutorial will follow the recommendations of @Peikert2019 and utilizes Git for version control, RMarkdown for dynamic document creation, Make for dependency tracking, and Docker for software management.
Their interplay is shown in Figure \@ref(fig:schematic).
At several points, we suggest other viable alternatives.

```{r schematic, eval = TRUE, echo = FALSE, fig.cap="Schematic illustration of the interplay of the four components (in dashed boxes) central to the reproducible workflow: version control (Git), dependency management (Make), software management (Docker), and dynamic document generation (R Markdown). Git tracks changes to the project over time. Make manages dependencies among the files. Docker provides a container in which the final report is built using dynamic document generation in R Markdown. Reproduced from @Peikert2019.  "}
# file gets downloaded in Makefile
knitr::include_graphics("images/nutshell.svg", auto_pdf = TRUE)
```

@@

<!-- researcher should begin early to enjoy most benefits -->
It can be tempting to leave considerations of reproducibility for when you finish a project.
<!-- weird sentence, I'll leave it as a placeholder-->
However, reproducibility is invaluable for collaboration---including with the version of yourself from a few weeks ago.
Therefore, we suggest starting early, ideally from the moment you decide to pursue a project to incorporate tools to help you archive reproducibility gradually.
The first step is to organize all files in one folder.
This way, you never lose anything, and you can easily share the files.
<!-- Hands-on: Introducing RProjects-->
RStudio's RProjects facilitate this concept of a project folder.
To create one, click on:

@@

@@

```{r, results='asis', echo=FALSE}
# in case we want to reuse this example later
markdown_example <- "
<!--this is a Markdown file -->
# Header
Normal text.
Important **word**.

@@

We will later see how this simplicity enables us to create many different document formats (i.e. Word, PDF, HTML) and incorporate code (i.e. R, Python, Julia).
Another advantage of Markdown is that you can easily version-control it with Git.
You may be familiar with Microsoft Words' "Track Changes"-Feature and version control is similar in its basic idea but more potent in its features.
Instead of tracking only one file, you can keep track of the whole project directory, and instead of losing all changes when you agree to them, the entire history is preserved transparently.
For example this paper has been going through `r length(git2r::commits())` `r link("https://github.com/aaronpeikert/repro-tutorial/commits/master", "iterations")`.
To activate Git in a given project, you can call:
<!-- Hands-on: Introduce Git -->

@@

And you could commit it in Git:
<!--I am leaning towards doing everything in R, not sure though-->
<!--For Git we maybe should use RStudio's Git pane-->

@@

And Git can show you what has changed, like in Figure \@ref(fig:idea-change).

```{r idea-change-screenshot, include=FALSE}
# unfortunately css selectors won't work for pdf so we'll use png instead
idea_change <- here::here("images", "idea-change.png")
if(!file.exists(idea_change)){
  webshot2::webshot("https://github.com/aaronpeikert/repro-tutorial/commit/5b3f4641ff542158184f85d458880c35e8f09f3c?diff=split",
                  file = idea_change,
                  selector = "div#diff-2f8beaa39e5c98706d8abb3361a8383776f8df8a195a21c0e6eeb25e5aa48f45", # only select file change
                  zoom = 2)# higher resolution
}
```

@@

<!--I am troubled by how we should recomend to learn git-->
We recommend to attend an online lecture or workshop to learn the ins and outs of Git.

@@

```{r, include=FALSE}
source(here::here("R", "simulation.R"))
```

There are probably few things as frustrating as realizing that the data that you gathered cannot answer the research question you had in mind.
A step to prevent unpleasant surprises is to simulate data.
Simulated data lets you verify that the data you expect to gather and the analysis you plan fit together.
Of course, one has to make many educated guesses to simulate data that closely matches your expectation.
These guesses strongly depend on the study you are planning, and there cannot be a one fits all solution.
However, the analytic strategy often gives you a hint about how to simulate data.
In the Machiavellism example, the analytic strategy is to use structural equation modelling (SEM).
To simulate data for SEM, one has to specify a model and then sample from a multivariate normal distribution following its implied covariance matrix.
The model parameters should mimic the Machiavellism subscale of the short dark triad (SD3).
As a starting point, we can use already published information.
@Jones2013 report loadings seen in Table \@ref(tab:loadings) and report a gender difference (at the manifest level) depending on the sample between `r min(cohend_jones_paulhus)` and `r max(cohend_jones_paulhus)`.

```{r loadings, echo=FALSE}
knitr::kable(data.frame(Item = 1:9,
                        Loading = loadings_jones_paulhus),
             caption = "Factor loadings from an Exploratory Structrural Equation Model as reported in Jones \\& Paulhus (2013)")
```

It is probably better to assume more pessimistic values.
This way, we make sure that our analytic strategy is viable even in non-optimal circumstances.
Therefore, we will assume that the loadings are 30% lower than reported and that the standardized mean difference is only 0.1.
These assumption can be simply translated into [`lavaan` [@lavaan]](https://lavaan.ugent.be)[^lavaan] model syntax:

[^lavaan]: https://lavaan.ugent.be

```{r, results='asis', echo=FALSE}
model_truth <- combine(
  measurement(
    "MACH",
    items("x", 9),
    loadings = round(loadings_jones_paulhus * 0.7, 2)
    ),
  intercepts("MACH", list(0, 0.1))
  )
cat("``` r",
    model_truth,
    "```",
    sep = "\n")
model <- combine(measurement(
  "MACH",
  items("x", 9)),
  intercepts("MACH", list(0, "NA")))
```

From which we can simulate data:

```{r, results='hide'}
simulated <- lavaan::simulateData(
  model_truth,
  sample.nobs = c(1000, 1000),
  meanstructure = TRUE,
  standardized = TRUE
)
```

And then fit the statistical model:

```{r}
fit <-
  lavaan::cfa(
    model,
    data = simulated,
    group = "group",
    meanstructure = TRUE,
    std.lv = TRUE
  ) 
```

But wait `lavaan` warns us that this model may not be identified.
We overlooked a crucial assumption of this model, that is measurement invariance.
This assumption is also baked into our simulation of the data.
It is probably more realistic to violate this assumption and plan our analysis accordingly.
Therefore, we simulate the data so that some of the items parameter differ among groups and then alter the analysis to assume partial measurement invariance.
While reconsidering our assumptions, it is also unlikely that the data is normally distributed.
Not only would the actual items be Likert-like and therefore categorical, but also are the in @Jones2013 reported means not in the middle of the scale, thus suggesting some skewness to the items.
@Rhemtulla2012 conclude that using Likert-like items is acceptable when they employ five or more categories.
Hence, we continue simulating continues data.
However, adding a little skewness to the items is easy and helps make the simulation more realistic.
Anyhow, remember that what we consider realistic for our case could be unfit for your case.
If you are satisfied with your assumptions it helps to write a function that expresses them, like this one:
<!--Hands-on: build simple functions tailored for your analysis-->

```{r generate}
```

Note however that this function is again composed of several functions, which you may inspect on `r link("https://github.com/aaronpeikert/repro-tutorial/blob/master/R/simulation.R", "GitHub")`.

With its help we can simulate 500 cases with a mean difference between groups of 0.1, where the intercepts of three random items vary by .3 and three loadings vary by 30% and five random items have a skew.

```{r}
simulated <- generate(n = 500,
                      diff = 0.1,
                      vary_intercepts = rep(0.3, 3),
                      vary_loadings = rep(0.3, 3),
                      skewness = rep(5, -0.5))
```

Note, that this certainly results in different numbers each time we run it.
To consistently get the same results it is good practice to use a seed.
The function `seeded_generate()` will always generate the same pseudo-random numbers, when given the same seed.

```{r}
seeded_generate <-  function(seed, ...)withr::with_seed(seed, generate(...))
simulated <- seeded_generate(
  seed = 1234,
  n = 100000,
  diff = 0.1,
  vary_intercepts = rep(0.3, 3),
  vary_loadings = rep(0.3, 3),
  skewness = rep(5,-0.5)
)
```

```{r}
lavaan::summary(analyze(simulated))
```

### Preregestration (optional)

@@

@@

@@

<!--Hands on: Add a license-->
```r
usethis::use_cc0_license()
```

@@

@@

@@

GitHub/GitLab are great as a shared information basis: sharing and editing files, discussing and distributing tasks via issues and documenting the project via wikis even before data collection starts.

<!--Hands-on: Introduce Github-->

@@

@@

### Internal Dependencies

@@

### External Dependencies

@@

#### renv (alternative)

#### Singularity (optional)

## Analyzing Data

<!-- Hands on: add data via automate() -->
<!-- Hands on: automatically download data via manually editing the Makefile -->

### Data Integrity (optional)

@@

### Anonymising Data (optional)

@@

## Peer Review (optional)

## Post Publication Review (optional)

## Archiving

@@

## Checklist
